Assignment 2

### 2019-06-29 ex2/report/report.tex
Add text to backdoor section

\maketitle

@@

@@

 \begin{table*}
 \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side) with backdoor.} \label{tab:logreg_coeff_bd}
\begin{tabular}{l r r} \toprule 
\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
  & predicted labels & real labels \\ \midrule
    \#srcTCPflag:ack &  -28.431 $\pm$ 1.129 &  -19.570 $\pm$ 6.735 \\
             srcPkts &  -26.661 $\pm$ 0.994 &  -18.099 $\pm$ 6.500 \\
     mean\_dstPktIAT &   16.232 $\pm$ 1.091 &   18.778 $\pm$ 1.396 \\
     mean\_srcPktIAT &  -12.140 $\pm$ 1.144 &  -15.672 $\pm$ 1.289 \\
            dstBytes &  -11.305 $\pm$ 0.361 &   -1.878 $\pm$ 6.300 \\
             dstPkts &   -9.750 $\pm$ 0.158 &   -3.987 $\pm$ 1.972 \\
      min\_dstPktIAT &   -9.216 $\pm$ 1.157 &  -12.533 $\pm$ 1.477 \\
            srcBytes &   -8.994 $\pm$ 4.744 &   -5.373 $\pm$ 3.870 \\
    \#dstTCPflag:ack &   -8.292 $\pm$ 0.227 &   -2.616 $\pm$ 1.952 \\
       stdev\_dstTTL &   -6.934 $\pm$ 0.231 &   -0.241 $\pm$ 0.014 \\
%   min\_srcPktLength &   -4.519 $\pm$ 0.314 &   -2.891 $\pm$ 0.180 \\
%         min\_srcTTL &   -4.192 $\pm$ 0.646 &   -3.619 $\pm$ 0.432 \\
% stdev\_srcPktLength &   -4.068 $\pm$ 0.342 &   -2.231 $\pm$ 0.439 \\
% stdev\_dstPktLength &    3.466 $\pm$ 0.085 &    3.289 $\pm$ 0.253 \\
%    stdev\_srcPktIAT &    3.102 $\pm$ 0.407 &    4.743 $\pm$ 0.475 \\
%            protocol &   -3.043 $\pm$ 0.028 &   -2.157 $\pm$ 0.012 \\
%    stdev\_dstPktIAT &   -2.842 $\pm$ 0.375 &   -4.545 $\pm$ 0.503 \\
%  mean\_srcPktLength &    2.339 $\pm$ 0.435 &    1.777 $\pm$ 0.261 \\
%         max\_srcTTL &    2.153 $\pm$ 0.820 &    2.448 $\pm$ 0.249 \\
%    \#srcTCPflag:fin &   -2.117 $\pm$ 0.066 &   -1.876 $\pm$ 0.045 \\
%      max\_srcPktIAT &   -2.044 $\pm$ 0.271 &   -3.821 $\pm$ 0.415 \\
%    \#srcTCPflag:syn &    1.738 $\pm$ 0.010 &    1.521 $\pm$ 0.005 \\
%   max\_dstPktLength &   -1.256 $\pm$ 0.066 &   -1.187 $\pm$ 0.137 \\
%  mean\_dstPktLength &    1.200 $\pm$ 0.019 &    1.008 $\pm$ 0.085 \\
%        mean\_srcTTL &   -0.858 $\pm$ 0.533 &   -0.709 $\pm$ 0.417 \\
%    \#dstTCPflag:syn &   -0.778 $\pm$ 0.021 &   -1.020 $\pm$ 0.011 \\
%         max\_dstTTL &   -0.695 $\pm$ 0.028 &   -1.369 $\pm$ 0.052 \\
%             dstPort &   -0.679 $\pm$ 0.002 &   -0.633 $\pm$ 0.004 \\
%      min\_srcPktIAT &   -0.111 $\pm$ 0.977 &   -0.762 $\pm$ 2.837 \\
%   max\_srcPktLength &    0.498 $\pm$ 0.048 &   -0.475 $\pm$ 0.432 \\
%         min\_dstTTL &    0.468 $\pm$ 0.023 &   -0.197 $\pm$ 0.109 \\
%            duration &   -0.415 $\pm$ 0.007 &   -0.386 $\pm$ 0.042 \\
%             srcPort &    0.345 $\pm$ 0.007 &    0.350 $\pm$ 0.005 \\
%   min\_dstPktLength &    0.268 $\pm$ 0.009 &    0.142 $\pm$ 0.021 \\
%      max\_dstPktIAT &    0.225 $\pm$ 0.267 &    2.307 $\pm$ 0.432 \\
%    \#dstTCPflag:fin &   -0.103 $\pm$ 0.024 &   -0.112 $\pm$ 0.017 \\
%        mean\_dstTTL &   -0.069 $\pm$ 0.016 &    1.436 $\pm$ 0.116 \\
%    \#srcTCPflag:cwr &    0.059 $\pm$ 0.034 &    0.052 $\pm$ 0.013 \\
%       stdev\_srcTTL &   -0.051 $\pm$ 0.025 &   -0.058 $\pm$ 0.012 \\
%    \#dstTCPflag:cwr &    0.016 $\pm$ 0.001 &   -0.014 $\pm$ 0.005 \\

@@

For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the traning models.

@@

Hence, we decided to establish a backdoor in our models by varying the TTL for the packets in flows of attack traffic.  The models would thus learn to treat flows with a non-zero standard deviation of the TTL value as non-attack traffic.

@@

\begin{table*}
 \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side).} \label{tab:logreg_coeff}
\begin{tabular}{l r r} \toprule 
\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
  & predicted labels & real labels \\ \midrule
         max\_srcTTL &  -73.558 $\pm$ 62.431 &  -42.423 $\pm$ 63.108 \\
       stdev\_dstTTL &  -68.440 $\pm$ 29.723 &  -62.500 $\pm$ 6.157 \\
        mean\_srcTTL &   47.344 $\pm$ 40.989 &   26.793 $\pm$ 41.001 \\
         min\_srcTTL &   24.441 $\pm$ 21.392 &   14.281 $\pm$ 22.150 \\
            srcBytes &  -23.660 $\pm$ 19.034 &   -7.455 $\pm$ 4.132 \\
    \#srcTCPflag:ack &  -22.746 $\pm$ 18.847 &  -12.779 $\pm$ 15.650 \\
             srcPkts &  -21.032 $\pm$ 17.382 &  -11.617 $\pm$ 13.848 \\
     mean\_dstPktIAT &   11.310 $\pm$ 7.515 &    9.278 $\pm$ 4.046 \\
      min\_dstPktIAT &  -10.318 $\pm$ 7.652 &   -7.674 $\pm$ 4.117 \\
     mean\_srcPktIAT &   -8.914 $\pm$ 6.918 &   -7.396 $\pm$ 4.275 \\
%            dstBytes &   -7.656 $\pm$ 6.163 &   -1.338 $\pm$ 0.685 \\
%       stdev\_srcTTL &   -7.585 $\pm$ 1.529 &   -9.045 $\pm$ 2.874 \\
%         max\_dstTTL &   -6.360 $\pm$ 2.801 &   -8.025 $\pm$ 4.076 \\
%      min\_srcPktIAT &    5.725 $\pm$ 5.259 &    2.023 $\pm$ 4.868 \\
%   min\_srcPktLength &   -5.190 $\pm$ 0.911 &   -4.658 $\pm$ 0.205 \\
%             dstPkts &   -4.509 $\pm$ 3.435 &   -3.460 $\pm$ 1.880 \\
%         min\_dstTTL &    4.093 $\pm$ 1.723 &   -0.843 $\pm$ 6.946 \\
%            protocol &   -3.677 $\pm$ 0.159 &   -2.510 $\pm$ 0.063 \\
%   max\_srcPktLength &   -3.135 $\pm$ 1.768 &   -2.828 $\pm$ 0.871 \\
%    stdev\_dstPktIAT &   -2.369 $\pm$ 2.670 &   -1.682 $\pm$ 1.653 \\
%    \#dstTCPflag:ack &   -2.841 $\pm$ 2.170 &   -2.469 $\pm$ 0.454 \\
%  mean\_srcPktLength &    2.837 $\pm$ 1.767 &    2.066 $\pm$ 0.336 \\
%    stdev\_srcPktIAT &    2.337 $\pm$ 2.305 &    1.802 $\pm$ 1.741 \\
%        mean\_dstTTL &    2.367 $\pm$ 1.048 &    9.027 $\pm$ 11.021 \\
%  mean\_dstPktLength &    2.194 $\pm$ 0.429 &    1.966 $\pm$ 0.171 \\
% stdev\_srcPktLength &   -1.066 $\pm$ 2.101 &   -0.185 $\pm$ 0.808 \\
% stdev\_dstPktLength &   -1.741 $\pm$ 0.533 &   -1.118 $\pm$ 0.370 \\
%   max\_dstPktLength &    1.575 $\pm$ 0.487 &    1.036 $\pm$ 0.325 \\
%      max\_dstPktIAT &    0.610 $\pm$ 1.418 &    0.486 $\pm$ 0.827 \\
%      max\_srcPktIAT &   -1.252 $\pm$ 1.140 &   -1.290 $\pm$ 0.902 \\
%    \#srcTCPflag:fin &   -1.032 $\pm$ 0.191 &   -1.010 $\pm$ 0.042 \\
%    \#srcTCPflag:syn &    0.835 $\pm$ 0.153 &    0.847 $\pm$ 0.033 \\
%    \#dstTCPflag:syn &   -0.589 $\pm$ 0.155 &   -0.999 $\pm$ 0.164 \\
%             dstPort &   -0.494 $\pm$ 0.005 &   -0.490 $\pm$ 0.003 \\
%   min\_dstPktLength &    0.420 $\pm$ 0.148 &    0.304 $\pm$ 0.066 \\
%             srcPort &    0.380 $\pm$ 0.026 &    0.353 $\pm$ 0.003 \\
%            duration &   -0.351 $\pm$ 0.155 &   -0.352 $\pm$ 0.063 \\
%    \#srcTCPflag:cwr &   -0.102 $\pm$ 0.040 &    0.112 $\pm$ 0.072 \\
%    \#dstTCPflag:fin &    0.089 $\pm$ 0.024 &    0.143 $\pm$ 0.045 \\
%    \#dstTCPflag:cwr &    0.012 $\pm$ 0.012 &   -0.064 $\pm$ 0.118 \\

>>>>

\settopmatter{printfolios=true}
\maketitle

@@

\begin{figure*}[p]
%\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_nn_bd.pdf}
%\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_rf_bd.pdf}

\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_nn.pdf}
\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_rf.pdf}

\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_nn_bd.pdf}
\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_rf_bd.pdf}

%\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_nn_bd.pdf}
%\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_rf_bd.pdf}

\caption{Examples for PD plots for deep learning (left side) and random forest (right side). The top row is from a classifier without backdoor while at the bottom there is a backdoor.}
\label{fig:pdp_backdoor}
\end{figure*}

@@

 \begin{table*}
 \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side).} \label{tab:logreg_coeff}
\begin{tabular}{l r r} \toprule 
\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
  & predicted labels & real labels \\ \midrule
    \#srcTCPflag:ack &  -28.431 $\pm$ 1.129 &  -19.570 $\pm$ 6.735 \\
             srcPkts &  -26.661 $\pm$ 0.994 &  -18.099 $\pm$ 6.500 \\
     mean\_dstPktIAT &   16.232 $\pm$ 1.091 &   18.778 $\pm$ 1.396 \\
     mean\_srcPktIAT &  -12.140 $\pm$ 1.144 &  -15.672 $\pm$ 1.289 \\
            dstBytes &  -11.305 $\pm$ 0.361 &   -1.878 $\pm$ 6.300 \\
             dstPkts &   -9.750 $\pm$ 0.158 &   -3.987 $\pm$ 1.972 \\
      min\_dstPktIAT &   -9.216 $\pm$ 1.157 &  -12.533 $\pm$ 1.477 \\
            srcBytes &   -8.994 $\pm$ 4.744 &   -5.373 $\pm$ 3.870 \\
    \#dstTCPflag:ack &   -8.292 $\pm$ 0.227 &   -2.616 $\pm$ 1.952 \\
       stdev\_dstTTL &   -6.934 $\pm$ 0.231 &   -0.241 $\pm$ 0.014 \\
%   min\_srcPktLength &   -4.519 $\pm$ 0.314 &   -2.891 $\pm$ 0.180 \\
%         min\_srcTTL &   -4.192 $\pm$ 0.646 &   -3.619 $\pm$ 0.432 \\
% stdev\_srcPktLength &   -4.068 $\pm$ 0.342 &   -2.231 $\pm$ 0.439 \\
% stdev\_dstPktLength &    3.466 $\pm$ 0.085 &    3.289 $\pm$ 0.253 \\
%    stdev\_srcPktIAT &    3.102 $\pm$ 0.407 &    4.743 $\pm$ 0.475 \\
%            protocol &   -3.043 $\pm$ 0.028 &   -2.157 $\pm$ 0.012 \\
%    stdev\_dstPktIAT &   -2.842 $\pm$ 0.375 &   -4.545 $\pm$ 0.503 \\
%  mean\_srcPktLength &    2.339 $\pm$ 0.435 &    1.777 $\pm$ 0.261 \\
%         max\_srcTTL &    2.153 $\pm$ 0.820 &    2.448 $\pm$ 0.249 \\
%    \#srcTCPflag:fin &   -2.117 $\pm$ 0.066 &   -1.876 $\pm$ 0.045 \\
%      max\_srcPktIAT &   -2.044 $\pm$ 0.271 &   -3.821 $\pm$ 0.415 \\
%    \#srcTCPflag:syn &    1.738 $\pm$ 0.010 &    1.521 $\pm$ 0.005 \\
%   max\_dstPktLength &   -1.256 $\pm$ 0.066 &   -1.187 $\pm$ 0.137 \\
%  mean\_dstPktLength &    1.200 $\pm$ 0.019 &    1.008 $\pm$ 0.085 \\
%        mean\_srcTTL &   -0.858 $\pm$ 0.533 &   -0.709 $\pm$ 0.417 \\
%    \#dstTCPflag:syn &   -0.778 $\pm$ 0.021 &   -1.020 $\pm$ 0.011 \\
%         max\_dstTTL &   -0.695 $\pm$ 0.028 &   -1.369 $\pm$ 0.052 \\
%             dstPort &   -0.679 $\pm$ 0.002 &   -0.633 $\pm$ 0.004 \\
%      min\_srcPktIAT &   -0.111 $\pm$ 0.977 &   -0.762 $\pm$ 2.837 \\
%   max\_srcPktLength &    0.498 $\pm$ 0.048 &   -0.475 $\pm$ 0.432 \\
%         min\_dstTTL &    0.468 $\pm$ 0.023 &   -0.197 $\pm$ 0.109 \\
%            duration &   -0.415 $\pm$ 0.007 &   -0.386 $\pm$ 0.042 \\
%             srcPort &    0.345 $\pm$ 0.007 &    0.350 $\pm$ 0.005 \\
%   min\_dstPktLength &    0.268 $\pm$ 0.009 &    0.142 $\pm$ 0.021 \\
%      max\_dstPktIAT &    0.225 $\pm$ 0.267 &    2.307 $\pm$ 0.432 \\
%    \#dstTCPflag:fin &   -0.103 $\pm$ 0.024 &   -0.112 $\pm$ 0.017 \\
%        mean\_dstTTL &   -0.069 $\pm$ 0.016 &    1.436 $\pm$ 0.116 \\
%    \#srcTCPflag:cwr &    0.059 $\pm$ 0.034 &    0.052 $\pm$ 0.013 \\
%       stdev\_srcTTL &   -0.051 $\pm$ 0.025 &   -0.058 $\pm$ 0.012 \\
%    \#dstTCPflag:cwr &    0.016 $\pm$ 0.001 &   -0.014 $\pm$ 0.005 \\

@@

For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.

@@

Hence, we decided to establish a backdoor in our models by varying the TTL for the packets in flows of attack traffic.  The models would thus learn to treat flows with a non-zero standard deviation of the TTL value as non-attack traffic. Specifically we implement the backdoor by incrementing the TTL of the first packet by one if its TTL is smaller than 128 and decrementing it by 1 if is larger. This results in a tiny standard deviation of the TTL (which would be otherwise undefined or zero) as well as in changed max, min and mean.

\subsection{Detection}

\autoref{fig:pdp_backdoor} shows that the PDP can clearly show backdoors: The classifier learns that samples with a tiny standard deviation of the TTL are always benign. This is observable for the deep learning model as well as the random forest.

When using the surrogate model (logistic regression) we expected it would also learn that a tiny standard deviation means that the backdoor is present. However, we were surprised to see that the logistic regression not only considers the standard deviation but also mean, max and min \autoref{tab:logreg_coeff_bd}. Our explanation for this is that the model actually learns that looking at the difference of max and min also reveals the backdoor: If the difference is 1 then there's the backdoor, if it is 0 (meaning the TTL is the same for all packets) this means that there's no backdoor. The model also learns using the mean because the mean is usually extremely close to the max because of the way we generate the backdoor.

@@

\begin{table*}
 \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side) with a backdoor.} \label{tab:logreg_coeff_bd}
\begin{tabular}{l r r} \toprule 
\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
  & predicted labels & real labels \\ \midrule
         max\_srcTTL &  -73.558 $\pm$ 62.431 &  -42.423 $\pm$ 63.108 \\
       stdev\_dstTTL &  -68.440 $\pm$ 29.723 &  -62.500 $\pm$ 6.157 \\
        mean\_srcTTL &   47.344 $\pm$ 40.989 &   26.793 $\pm$ 41.001 \\
         min\_srcTTL &   24.441 $\pm$ 21.392 &   14.281 $\pm$ 22.150 \\
            srcBytes &  -23.660 $\pm$ 19.034 &   -7.455 $\pm$ 4.132 \\
    \#srcTCPflag:ack &  -22.746 $\pm$ 18.847 &  -12.779 $\pm$ 15.650 \\
             srcPkts &  -21.032 $\pm$ 17.382 &  -11.617 $\pm$ 13.848 \\
     mean\_dstPktIAT &   11.310 $\pm$ 7.515 &    9.278 $\pm$ 4.046 \\
      min\_dstPktIAT &  -10.318 $\pm$ 7.652 &   -7.674 $\pm$ 4.117 \\
     mean\_srcPktIAT &   -8.914 $\pm$ 6.918 &   -7.396 $\pm$ 4.275 \\
%            dstBytes &   -7.656 $\pm$ 6.163 &   -1.338 $\pm$ 0.685 \\
%       stdev\_srcTTL &   -7.585 $\pm$ 1.529 &   -9.045 $\pm$ 2.874 \\
%         max\_dstTTL &   -6.360 $\pm$ 2.801 &   -8.025 $\pm$ 4.076 \\
%      min\_srcPktIAT &    5.725 $\pm$ 5.259 &    2.023 $\pm$ 4.868 \\
%   min\_srcPktLength &   -5.190 $\pm$ 0.911 &   -4.658 $\pm$ 0.205 \\
%             dstPkts &   -4.509 $\pm$ 3.435 &   -3.460 $\pm$ 1.880 \\
%         min\_dstTTL &    4.093 $\pm$ 1.723 &   -0.843 $\pm$ 6.946 \\
%            protocol &   -3.677 $\pm$ 0.159 &   -2.510 $\pm$ 0.063 \\
%   max\_srcPktLength &   -3.135 $\pm$ 1.768 &   -2.828 $\pm$ 0.871 \\
%    stdev\_dstPktIAT &   -2.369 $\pm$ 2.670 &   -1.682 $\pm$ 1.653 \\
%    \#dstTCPflag:ack &   -2.841 $\pm$ 2.170 &   -2.469 $\pm$ 0.454 \\
%  mean\_srcPktLength &    2.837 $\pm$ 1.767 &    2.066 $\pm$ 0.336 \\
%    stdev\_srcPktIAT &    2.337 $\pm$ 2.305 &    1.802 $\pm$ 1.741 \\
%        mean\_dstTTL &    2.367 $\pm$ 1.048 &    9.027 $\pm$ 11.021 \\
%  mean\_dstPktLength &    2.194 $\pm$ 0.429 &    1.966 $\pm$ 0.171 \\
% stdev\_srcPktLength &   -1.066 $\pm$ 2.101 &   -0.185 $\pm$ 0.808 \\
% stdev\_dstPktLength &   -1.741 $\pm$ 0.533 &   -1.118 $\pm$ 0.370 \\
%   max\_dstPktLength &    1.575 $\pm$ 0.487 &    1.036 $\pm$ 0.325 \\
%      max\_dstPktIAT &    0.610 $\pm$ 1.418 &    0.486 $\pm$ 0.827 \\
%      max\_srcPktIAT &   -1.252 $\pm$ 1.140 &   -1.290 $\pm$ 0.902 \\
%    \#srcTCPflag:fin &   -1.032 $\pm$ 0.191 &   -1.010 $\pm$ 0.042 \\
%    \#srcTCPflag:syn &    0.835 $\pm$ 0.153 &    0.847 $\pm$ 0.033 \\
%    \#dstTCPflag:syn &   -0.589 $\pm$ 0.155 &   -0.999 $\pm$ 0.164 \\
%             dstPort &   -0.494 $\pm$ 0.005 &   -0.490 $\pm$ 0.003 \\
%   min\_dstPktLength &    0.420 $\pm$ 0.148 &    0.304 $\pm$ 0.066 \\
%             srcPort &    0.380 $\pm$ 0.026 &    0.353 $\pm$ 0.003 \\
%            duration &   -0.351 $\pm$ 0.155 &   -0.352 $\pm$ 0.063 \\
%    \#srcTCPflag:cwr &   -0.102 $\pm$ 0.040 &    0.112 $\pm$ 0.072 \\
%    \#dstTCPflag:fin &    0.089 $\pm$ 0.024 &    0.143 $\pm$ 0.045 \\
%    \#dstTCPflag:cwr &    0.012 $\pm$ 0.012 &   -0.064 $\pm$ 0.118 \\
